\documentclass{article}

\usepackage{enumitem}
\setlist[enumerate, 2]{label=\alph*.}

\usepackage{siunitx}

\title{EE 450 Homework 1}
\author{Shangning Xu}

\begin{document}

\maketitle

\begin{enumerate}
    \item[P4.] \begin{enumerate}
        \item We can have 4 connections between each pair of adjacent switches, giving a total of 16 connections.
        \item 8, with 4 connections passing through the lower-left switch and 4 through upper-right.
        \item Yes. We can route 2 connections between A and C through B and the other 2 through D, and route connections between B and D through A and C.
    \end{enumerate}

    \item[P5.] \begin{enumerate}
        \item The textbook has shown that the transmission delay is \SI{2}{min} through each tollbooth, and the caravan propagates for a total of \SI{150}{km}, so the end-to-end delay is
        \[
            \SI{150}{km} / (\SI{100}{km/h}) + \SI{2}{min} \cdot 3 = \SI{96}{min}.
        \]
        \item With only eight cars, the transmission delay is now $8 / (\SI{5}{/min}) = \SI{1.6}{min}$, so the end-to-end delay is
        \[
            \SI{150}{km} / (\SI{100}{km/h}) + \SI{1.6}{min} \cdot 3 = \SI{94.8}{min}.
        \]
    \end{enumerate}

    \item[P6.] \begin{enumerate}
        \item $d_\textrm{prop} = m/s$ seconds
        \item $d_\textrm{trans} = L/R$ seconds
        \item $(L/R + m/s)$ seconds
        \item It is at A's end of the link.
        \item It is still in the link.
        \item The first bit has reached Host B.
        \item We have
        \[
            \frac{m}{2.5 \cdot 10^8} = \frac{120\textrm{ bits}}{56\textrm{ kbps}}.
        \]
        Solving for $m$, we get $m \approx 5.4 \cdot 10^5$.
    \end{enumerate}

    \item[P7.] The end-to-end delay for a packet is $\SI{56}{B} / \SI{2}{Mbps} + \SI{10}{ms} = \SI{10.224}{ms}$. However, bits in a packet are not created at the same time. The first bit in the packet is created $\SI{56}{B} / \SI{64}{kbps} = \SI{7}{ms}$ before the last bit. Therefore, 10.224--17.224 ms elapses from the time a bit is created until it is decoded.

    \item[P10.] The end-to-end delay is
    \[
        2d_\textrm{proc} + \sum_{i = 1}^3 \left(\frac{L}{R_i} + \frac{d_i}{s_i}\right).
    \]
    Substituting actual values into the formula, we obtain 64 ms for the end-to-end delay.

    \item[P12.] The queuing delay is $\SI{1500}{B} \cdot 4.5 / \SI{2}{Mbps} = \SI{27}{ms}$. Similarly, the queuing delay for the general case is $(nL + L - x)/R$.
    
    \item[P20.] $\min\{R_s, R_c, R/M\}$
    
    \item[P21.] If the server can only use one path, the maximum throughput is
    \[
        \max_{k = 1, \dots, M} \min_{i = 1, \dots, N} R_i^k.
    \]
    If the server can use all $M$ paths, the maximum throughput is
    \[
        \sum_{k = 1}^M \min_{i = 1, \dots, N} R_i^k.
    \]

    \item[P25.] \begin{enumerate}
        \item $R \cdot d_\textrm{prop} = \SI{2}{Mbps} \cdot \SI{20000}{km} / (\SI{2.5d8}{m/s}) = \SI{160}{kb}.$
        \item The number of bits that will be in the link is at its maximum when both B is receiving and A is transmitting. Because the first bit and the last bit in the link is $d_\textrm{prop}$ time apart when transmitted into the link, there are at most $R \cdot d_\textrm{prop} = \SI{160}{kb}$ in the link.
        \item The maximum possible number of bits in the link at any given time.
        \item The width of a bit is $\SI{20000}{km} / \num{160000} = \SI{125}{m}$, longer than an American football field at 360 feet (110 m).
        \item $m / (Rm/s) = s/R$
    \end{enumerate}

    \item[P28.] \begin{enumerate}
        \item The total time it takes to send the file is
        \[
            \SI{800000}{b} / \SI{2}{Mbps} + \SI{20000}{km} / (\SI{2.5d8}{m/s}) = \SI{0.48}{s}.
        \]
        \item Because the transmission time of the acknowledgement packet is ignored, we only need to consider its propagation delay. The new total time it takes to send the file is now
        \[
            20(\SI{40000}{b} / \SI{2}{Mbps} + \SI{20000}{km} / (\SI{2.5d8}{m/s}) \cdot 2) = \SI{3.6}{s}.
        \]
        \item After the file is broken up into packets, transmission time is now 7.5 times as much as before, because propagation of file bits across different packets no longer overlaps and there is an extra propagation delay for each packet due to the acknowledgement packet, contributing to large increase in total propagation time.
    \end{enumerate}
\end{enumerate}

\end{document}
